\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{parskip}

\lstset{
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue},
  commentstyle=\color{gray},
  frame=single,
  breaklines=true,
}

\newcommand{\code}[1]{\texttt{\detokenize{#1}}}

\title{Design Document: Align Mode for WhyRel}
\author{WhyRel Project}
\date{\today}

\begin{document}
\maketitle

\tableofcontents
\newpage

% ============================================================
\section{About}
% ============================================================

\cite{banerjee-toplas-2022} provide a framework for verifying 2-safety which involves
constructing 2-products from unary while-programs with heaps called
\emph{bicommands} (Bicoms). This is translated to WhyML while
respecting the underlying relational region logic and then verified using Why3. The current implementation of the
framework, i.e. WhyRel, applies to 2-liveness verification as well by extending the language
of biprograms. In general, Bicoms pair the execution steps of two unary methods,
together with annotations that serve as the proof skeleton for relational
verification. The choice of pairing determines the intermediate annotation needed to verify the specification. Thus an alignment that yields simple intermediate assertions is needed. Manual construction of alignment i.e. bicommands is tedious, error-prone, and
time-consuming; moreover, an alignment suited to the relational
specification often requires several iterations to discover.  

This document presents \emph{Align mode}, a workflow within WhyRel that helps construct \emph{bimethods} which are bicommands over unary methods. The steps are as follows:

\begin{enumerate}
  \item CLI invocation (\S\ref{sec:use}): Receives two unary methods along with a relational specification from command line.
  \item Default Alignment: Computes a \emph{sequential alignment}, the bicommand $(C \mid C')$, of two unary method bodies $C$, $C'$
  \item Serving: A local HTTP API (and a browser UI) for inspection; and
  read-only (text/HTML) or interactive (the API above).
  \item Optimizing: User manipulates the alignment using \emph{rewrite} rules (\S\ref{sec:rewrite})  and suggest  annotations. 
  \item Export: The bimethod can then be exported as a .rl text file.
\end{enumerate}

An overview of the codebase is in \S\ref{sec:code}.

% ============================================================
\section{Usage}
\label{sec:use}
% ============================================================

Align mode is invoked as:
\begin{lstlisting}
whyrel -align -l <LMod> <LMeth> -r <RMod> <RMeth> files...
       [-o <file>] [-i] [-port <n>]
       [-rpre <rformula>] [-rpost <rformula>]
\end{lstlisting}

\begin{description}
  \item[\code{-l} / \code{-r}] left/right module and method to align.
  \item[\code{-o <file>}] output file for export (default
        \code{<stem>_align_output.rl}).
  \item[\code{-i}] start the \emph{interactive} rewriting server (\S\ref{sec:interactive}).
  \item[\code{-port <n>}] port for any of the servers (default \code{8080}).
  \item[\code{-rpre} / \code{-rpost}] take single relational formulas in WhyRel syntax 
      (e.g.\ \code{"Agree x"}, \code{"result =:= result"}, \code{"Both(n > 0)"}). More in (\S\ref{sec:spec}).
\end{description}

Example: cd to \code{examples/all_all/factorial} and run \code{whyrel -align -i -l M0 fact -r M1 fact -port 8083 prog.rl}


The specified methods and specifications are parsed and type-checked. An http server is setup at localhost:8083 serving a bicommand over M0-fact
and M1-fact methods in a default alignment. A browser ui is available at
\code{/ui} endpoint. Without \code{-i}, the default aligned bicommand is written to
the output file and printed to stdout, then the process exits.  

Since the interaction is mediated via http server, one can direct an LLM to this
section and have it figure out the rewrite functionality by reading the help
descriptions provided.

% ============================================================
\section{Relational specifications}
\label{sec:spec}
% ============================================================
\code{-rpre}/\code{-rpost} are parsed using a new parser start symbol \code{rformula_top} added in \code{parser.mly} and a new function
\code{parse_rformula_string} in \code{Astutil}. The formula is then type-checked with \code{Typing.tc_rformula}
against a \emph{restricted} \code{bi_tenv} built by \code{restricted_tenv}:
starting from \code{initial_tenv} (only \code{alloc}), over the program's class
table, extended with exactly the method parameters ---and \code{result} for the
postcondition only. Any reference outside the arguments (or \code{result})
therefore fails type-checking with an ``unknown identifier'' error. 

\code{-rpre}/\code{-rpost} can be accessed via \code{/spec} endpoint. For
\code{/export}: when \code{-rpre}/\code{-rpost} are supplied they become the
exported \code{requires}/\code{ensures}; otherwise \code{bimodule_skeleton}
derives relational defaults from the two unary specs (each side's pre/post
lifted with \code{Rleft}/\code{Rright}, plus parameter/result \code{=:=}
agreements).

% ============================================================
\section{Default alignment}
% ============================================================

The default alignment is the sequential alignment and is computed in \code{Auto.compose_sequentially}. It
 produces the \emph{sequential} bicommand
$(C \mid C')$: it runs all of the left program and then all of the right
program. This is the least aligned adequate product.
It is always projection-valid, and it is the starting point from which the
weaving rewrites introduce synchronisation.

% ============================================================
\section{The rewrite algebra (\texttt{rewrites.ml})}
\label{sec:rewrite}
% ============================================================

A rewrite is a partial, projection-preserving transformation on a bicommand:
\begin{lstlisting}
type rewrite = bicommand -> bicommand option  
\end{lstlisting}

\subsection{Forward weaving (coarse $\to$ fine)}


The rewrites follow the \emph{weaving relation} in ~\cite{banerjee-toplas-2022} (our version of biprograms have been extended with \code{assert}/\code{assume} bicommands). Every
rewrite preserves the projections of the bicommand, so any sequence of rewrites
applied to the default alignment is again an alignment of the same two unary
methods.


\begin{lstlisting}
weave_vardecl : (Var x | Var y in (C|C'))  -->  Var x|y in (C|C')
weave_seq     : (C1;C2 | C1';C2')          -->  (C1|C1') ; (C2|C2')
weave_sync    : (A | A)                    -->  |_ A _|   
weave_if      : (if E .. | if E' ..)       -->  if E|E' 
                                                then (..|..) 
                                                else (..|..)
weave_if4     : (if E .. | if E' ..)       -->  Biif4 with all four branch 
                                                      pairings
weave_while   : (while E .. | while E' ..) -->  while E|E' . <false|false> 
                                                do (..|..)
\end{lstlisting}

\begin{itemize}
  \item \code{weave_seq} Applied repeatedly down a right-nested sequence it peels the two programs
        apart one statement at a time.
  \item \code{weave_sync} fires only when the two atomic commands are
        structurally equal (the test used by \code{Annot.eqv_command}).
  \item \code{weave_if} produces the two-way \code{Biif} (lockstep); it is
        projection-valid always but \emph{adequate} only when the guards agree
        --- that becomes a proof obligation. \code{weave_if4} produces the
        four-way \code{Biif4}, which is adequate unconditionally but de-aligns;
        it is offered in \code{/suggest} but not used by \code{auto_align}.
  \item \code{weave_while} generates the \emph{lockstep} loop (both alignment
        guards \code{false}) and carries the unary loop invariants across
        (\S\ref{sec:invariants}).
 \item \code{weave_while_guard ag} build biwhiles with custom alignment guards ag.
\end{itemize}

\subsection{Inverse weaving (fine $\to$ coarse)}

\begin{lstlisting}
unweave_seq : (C1|C1') ; (C2|C2') -->  (C1;C2 | C1';C2')
unsync      : |_ A _|             -->  (A | A)
\end{lstlisting}

\subsection{Skip rules (one side at a time)}

These split a \code{Bisplit} into a sequence of one-sided steps:

\begin{lstlisting}
skip_split  : (C | C')              -->  (C | skip) ; (skip | C')
left_first  : (a;rest | C')         -->  (a | skip) ; (rest | C')
right_first : (C | a;rest)          -->  (skip | a) ; (C | rest)
lrc         : (C | skip);(skip | D) -->  (skip | D) ; (C | skip)
\end{lstlisting}

\code{skip_split} only fires when neither side is already
\code{skip} (so it cannot loop). \code{lrc} (left/right commute) from bikat \cite{antonopoulose-2023-popl} swaps an
adjacent left-only step and right-only step --- the two touch disjoint sides, so
the order is interchangeable and each side's projection is unchanged.


\subsection{Carrying and suggesting invariants}
\label{sec:invariants}

Weaving a loop must not discard the invariants the unary loops were verified
with. \code{lift_unary_invariants} converts them: an invariant present
(syntactically) on \emph{both} loops becomes \code{Rboth}, a left-only one
\code{Rleft}, a right-only one \code{Rright}; \code{biwhile_spec_of} also pairs
the two frames. This is projection-safe --- \code{projl}/\code{projr} recover
each side's invariants, the other side's projecting to \code{Ftrue} --- and both
\code{weave_while} and \code{weave_while_guard} use it, so weaving a loop no
longer throws the invariants away.

On top of the carried invariants the tool suggests \emph{relational} ones.
\code{coupling_candidates} builds agreement invariants \code{v =:= v} for each
variable occurring in both loops (collected from the guards and bodies by
\code{evars_of_exp} / \code{evars_of_command}). 

\code{add_invariant rf} adds a relational invariant \code{rf} to a
\code{Biwhile}. The interactive layer feeds it either a chosen candidate or an
arbitrary, user supplied formula (\S\ref{sec:invsuggest}).

\subsection{Focusing}

A rewrite applies at a sub-term selected by a path of child indices:
\begin{lstlisting}
type path = int list   (* e.g. [0;1]; [] = root *)
child_count, get_child, with_child   (* the zipper primitives *)
apply_at  : path -> rewrite -> bicommand -> bicommand option
subterm_at: path -> bicommand -> bicommand option
scope_at  : path -> bicommand -> varbind list * varbind list  (* locals in scope at the focus *)
suggest_at: path -> bicommand -> (string * bicommand) list  (* applicable rewrites + results *)
\end{lstlisting}

\code{scope_at} collects the \code{Bivardecl} bindings passed through on the way
to a focus (split into left/right); with the method parameters it gives the
typing context for an invariant at a loop (\S\ref{sec:invsuggest}).

\code{all_rewrites} is the named registry used by \code{suggest_at}
(\code{weave_vardecl}, \code{weave_seq}, \code{weave_sync}, \code{weave_if},
\code{weave_if4}, \code{weave_while}, \code{unweave_seq}, \code{unsync},
\code{skip_split}, \code{left_first}, \code{right_first}, \code{lrc}). The
parameterised
rewrites (\code{weave_while_guard}, \code{add_invariant}, \code{remove_invariant},
\code{change_ag}) are \emph{not} in the registry --- they need an argument and are
dispatched by name in \code{interactive.ml}.

\subsection{Auto Alignment}

\code{auto_align} repeatedly applies the forward weaving rules, descending into
the structure, to turn the sequential default into the finest \emph{lockstep}
alignment the shared structure allows. Statements that do not line up are left
as a \code{Bisplit}. It is only a suggestion; the user refines it.



% ============================================================
\section{Interactive server (\texttt{interactive.ml})}
\label{sec:interactive}
% ============================================================

With \code{-i}, \code{Interactive.run} builds the default alignment, checks the
spec, and serves a stateful API on the chosen port. State is a mutable
\code{current} bicommand plus two stacks: \code{history} (undo) and \code{future}
(redo).

\subsection{HTTP API}

\begin{lstlisting}
GET  /                        help text
GET  /ui                      single-page browser UI
GET  /bicom                   current alignment (canonical text)
GET  /bicom-tree              current alignment as JSON [{lineno,path,text}]
GET  /spec                    relational pre/postcondition (read-only)
GET  /history                 undo/redo availability and stack depths (JSON)
GET  /rules                   rewrite-registry rule names (JSON)
GET  /suggest                 every applicable rewrite, all foci (JSON)
GET  /suggest?path=0.1        applicable rewrites at one focus (JSON)
GET  /invariants?path=P       carried invariants + invariant candidates at P (JSON)
POST /rewrite?rule=R&path=P                 apply rewrite-registry rule R at focus P
POST /rewrite?rule=weave_while&guard_left=L&guard_right=R&path=P
                                weave a loop with custom alignment guard <L|R>
                                (empty guards => lockstep default <false|false>)
POST /rewrite?rule=change_ag&guard_left=L&guard_right=R&path=P
                                replace a woven loop's alignment guard pair
POST /rewrite?rule=add_invariant&formula=R&path=P    add relational invariant R
POST /rewrite?rule=remove_invariant&formula=R&path=P drop carried invariant R
POST /undo  /  POST /redo                   undo / redo (verb endpoints only)
POST /auto                    greedily weave toward lockstep
POST /reset                   back to the sequential default
POST /export                  write a .rl bimodule skeleton to the output file
POST /verify?t=N&theory=T     translate the skeleton and discharge its VCs with
                                the prover portfolio, asynchronously; returns 202
                                immediately (t = per-goal seconds, default 30;
                                theory defaults to the bimodule)
GET  /verify/status           running | idle | the finished run's result
                                (driver JSON, or error text)
POST /verify/abort            kill the running verification
\end{lstlisting}

A \code{path} is a dot-separated list of child indices (\code{0.1}); empty is the
root. Mutating endpoints return the updated alignment as text; \code{409} means
the rule did not apply at the focus, \code{400} a malformed request.

The verbs are enforced, not decorative: every state-changing endpoint
(\code{/rewrite}, \code{/undo}, \code{/redo}, \code{/auto}, \code{/reset},
\code{/export}, \code{/verify}, \code{/verify/abort}) rejects a non-\code{POST}
request with \code{405}. Read-only endpoints remain \code{GET}-able. This
matters because the server is a plain unauthenticated localhost service: a
browser prefetch, a link preview, or a stray \code{GET} of \code{/verify} would
otherwise silently start a prover run --- or, at \code{/reset}, discard the
user's alignment.

\code{/bicom} is the canonical \code{Pretty.pp_bicommand} text; \code{/bicom-tree}
is the per-line JSON the UI renders (each line tagged with the tree path it
belongs to, for click-to-focus). They carry the same alignment but
\code{/bicom-tree} rebuilds loop/if headers itself, so the two are not
byte-identical; \code{/bicom} is kept as the source of the exact text.

\subsection{Undo / redo}

Every successful rewrite \code{snapshot}s
the previous state onto \code{history} and clears \code{future}; \code{/undo}
moves the top of \code{history} to \code{future} and back to \code{current};
\code{/redo} is the mirror image. Undo/redo span \emph{rewrite} steps only:
\code{/auto} and \code{/reset} clear both stacks.

Undo/redo are kept \emph{distinct} from rewrites: they are session-history
operations, not tree transformations, so they live only at the verb endpoints
\code{/undo} and \code{/redo}. \code{/rewrite} applies the
\code{Rewrites.all_rewrites} registry exclusively --- there are no
\code{rule=undo}/\code{rule=redo} aliases, and \code{/rules} lists only real
rewrites. \code{/history} reports availability and stack depths.

\subsection{Biwhile alignment guards}
\label{sec:guards}

Loop weaving with a custom alignment guard is done directly via:
\begin{lstlisting}
POST /rewrite?rule=weave_while&guard_left=L&guard_right=R&path=P
\end{lstlisting}
where \code{L} and \code{R} are arbitrary relational formulas (rformulas) for
the left-side and right-side alignment guards. Each
side is parsed with \code{parse_rformula_string} and type-checked in the
loop-local \code{bi_tenv} (method parameters, \code{result}, and all locals
in scope at the focus) before the rewrite is applied.

The browser UI exposes the two guard fields as \emph{Left Guard} and
\emph{Right Guard}; leaving both empty applies the lockstep default
(\code{Both (false)} on each side).

\paragraph{Changing a woven loop's guard.} Once a loop is a \code{Biwhile},
its alignment guard pair can be replaced in place with
\begin{lstlisting}
POST /rewrite?rule=change_ag&guard_left=L&guard_right=R&path=P
\end{lstlisting}
\code{change_ag} takes the \emph{pair} \code{(L, R)} (parsed/type-checked by the
same \code{parse_custom_guard}), since an alignment guard is a left/right pair.

\subsection{Biwhile Invariants}
\label{sec:invsuggest}

\code{/invariants?path=P} reports, for a loop focus, the invariants already
carried (\S\ref{sec:invariants}) and the relational candidates (for
\emph{discovery}). Invariants are added and removed by formula:
\begin{itemize}
  \item \code{POST /rewrite?rule=add_invariant&formula=R&path=P} --- add an
        arbitrary, user supplied relational invariant \code{R}; and
  \item \code{POST /rewrite?rule=remove_invariant&formula=R&path=P} --- drop a
        carried invariant matching \code{R}.
\end{itemize}

Both forms require the loop to be woven (a \code{Biwhile}) first.

\subsection{Verification (\texttt{/verify})}
\label{sec:verify}

\code{/verify} closes the loop from alignment to proof: it takes the alignment
currently in the session, runs it through the same skeleton generation as
\code{/export}, translates it to WhyML, and discharges the resulting VCs with the
prover portfolio --- without leaving the browser. The user edits an alignment,
presses \emph{Verify}, and sees which VCs survive.

\paragraph{The pipeline.} \code{spawn_verify} writes the bimodule skeleton to a
temporary \code{.rl}, then runs a two-stage shell command:
\begin{lstlisting}
whyrel <program files> <tmp.rl> -o <tmp.mlw>  || exit 90
python3 tools/whyrel_prove.py <tmp.mlw> -L <dirs> -T <theory> -t N --json <tmp.json>
\end{lstlisting}
The library flags are the directories of the program files plus the stdlib. The
sentinel exit code \code{90} (\code{translate_failed_rc}) distinguishes a
\emph{translation} failure from the driver's own exit codes, which are \code{0}
(all goals proved) and \code{1} (some goals open) --- both of which carry a valid
JSON result. Without the sentinel the two cases are indistinguishable, since a
failed translation also exits nonzero.

\paragraph{Asynchrony.} The server is single-threaded (\code{Lwt}), so a
synchronous prover run would block the accept loop --- and in particular would
make \code{/verify/abort} unreachable, since the abort request could not be
received while the handler it must interrupt is still running. Verification is
therefore spawned as a detached child and polled. Session state is
\begin{lstlisting}
[ `Idle | `Running of verify_run | `Done of int * string * string ]
\end{lstlisting}
where \code{verify_run} records the child pid and the temporary JSON and log
paths. \code{/verify} returns \code{202 Accepted} the moment the child is
spawned; \code{/verify/status} calls \code{waitpid} with \code{WNOHANG},
transitions \code{Running} $\to$ \code{Done} exactly once, and serves the cached
result thereafter. Starting a run while one is in flight is a \code{409}, not a
second process. On \code{ECHILD} --- the child already reaped, e.g.\ across a
restart --- status falls back to reading the artifacts on disk.

\paragraph{Aborting.} The child is launched under \code{setsid}, so it heads its
own process group. \code{/verify/abort} signals the \emph{group}
(\code{kill(-pid)}), not just the shell: the pipeline spawns \code{whyrel},
\code{python3}, and in turn several prover processes, and signalling only the
immediate child would orphan the provers, leaving them to burn CPU after the
user has already moved on.

\paragraph{Status codes.} \code{200} with the driver's JSON for a completed run
(exit \code{0} or \code{1}); \code{409} plus the captured log if translation
failed; \code{500} plus the log for any other driver failure; \code{400} if the
CLI specs are incomplete, in which case verification is blocked for the same
reason \code{/export} is.

\paragraph{Prover configuration.} \code{/verify} invokes
\code{whyrel_prove.py} with its defaults --- the three-prover portfolio
(\code{alt-ergo}, \code{cvc5}, \code{z3}) under \code{split_vc}, at \code{-t 30}.
It does not currently expose the driver's \code{--emulate-auto3} flag, so the
push-button path runs the flat configuration rather than the driver's
partial emulation of Why3's \code{Auto\_level\_3} (four provers under
\code{split\_all\_full}; note this omits the strategy's
\code{introduce\_premises} and \code{inline\_goal} steps, so it is not
equivalent to the strategy itself). Threading a strategy selector through
\code{/verify} and the UI is an obvious next step (\S\ref{sec:open}).

% ============================================================
\section{HTTP transport and browser UI}
% ============================================================

\subsection{HTTP transport (\texttt{http\_server.ml})}

\code{http_server.ml} contains \code{start_dispatch_server}: a generic request-dispatching server
used by the interactive session.  The caller supplies a handler
\code{~meth ~path ~query -> (status, content-type, body)}; the server parses the
request line (\code{parse_request}, \code{query_params}), calls the handler, and
formats the HTTP response (\code{http_response}).  Handler exceptions become
\code{500}.

The interactive API passes all parameters in the query string, so request bodies
are never parsed.  \code{query_params} percent-decodes keys and values
(\code{url_decode}), allowing arguments such as a relational invariant passed to
\code{add_invariant} to contain spaces and punctuation.  The server runs through
\code{run_accept_loop}, which installs SIGINT/SIGTERM handlers and closes the
listening socket for a clean shutdown.

\subsection{Browser UI (\texttt{ui.ml})}

\code{ui.ml} contains the rendering function \code{build_interactive_html},
which generates the full single-page application served at \code{/ui}.  The page
drives the API directly from the browser: it renders the alignment from
\code{/bicom-tree} (each line tagged with its tree path, so a line number is
click-to-focus), loads suggestions from \code{/suggest}, applies rewrites at a
focus (filling in path, rule, guards, and invariant), and runs undo/redo,
auto-align, reset, and export. Rendering happens client-side; the server only
ships data (see \code{docs/align.md} for the request/response cycle).

The page also drives verification (\S\ref{sec:verify}): \emph{Verify} posts to
\code{/verify} with the timeout from a small input box, then polls
\code{/verify/status} every two seconds, swapping the button pair
(\code{verifyUiRunning}) so that \emph{Abort} is live exactly while a run is in
flight. \code{renderVerify} summarises the driver JSON --- per-prover counts, the
union, and each open goal with its explanation --- into the verification panel.

\code{interactive.ml} calls \code{build_interactive_html} when responding to
\code{/ui}; it is never called from \code{http_server.ml}.

% ============================================================
\section{Export and Pretty Printing}
% ============================================================

\code{/export} writes to the
output file the current woven bicommand wrapped in a \code{bimodule} stub. The biinterface and method signatures are left for the user
to fill in. \code{Pretty.pp_bicommand} renders bicommands.


% ============================================================
\section{Design decisions and open questions}
\label{sec:open}
% ============================================================

\begin{enumerate}
  \item \textbf{Projection correctness as invariant.} Every rewrite preserves
        $\pi_L(B) \equiv L$ and $\pi_R(B) \equiv R$. Rewrites do \emph{not}
        guarantee fault-freedom/adequacy (e.g.\ \code{weave_if} bets on guard
        agreement); that remains the prover's responsibility, as in the weaving
        discipline.

  \item \textbf{configurable port.} Servers take \code{-port}
        (default 8080)

  \item \textbf{state and undo.} The interactive server holds a single
        \code{current} reference plus \code{history}/\code{future} stacks; focus
        is an explicit path argument rather than a zipper held in state.

  \item \textbf{Spec environment.} The scope check rejects everything but
        arguments/result because the restricted environment omits user predicate
        signatures. Seeding it from the modules' predicate declarations would let
        specs mention user-defined relational predicates.

  \item \textbf{Export fidelity.} The exported skeleton has placeholder method
        signatures and no biinterface. Deriving the biinterface (cf.\
        \code{pretrans/derive_biinterface.ml}) and the real signatures would let
        the result feed straight back into the WhyRel pipeline.

\end{enumerate}



% ============================================================
\section{Code overview}
\label{sec:code}
% ============================================================

\begin{description}
  \item[\code{align.ml}] is the entry point into the workflow.
  \item[\code{auto.ml}] has sequential composition function.
  \item[\code{rewrites.ml}] the rewrite algebra: weaving/unweaving and skip
        rules, invariant/guard rewrites, coupling-candidate generation, focus
        paths, and the \code{auto_align} driver.
  \item[\code{interactive.ml}] the stateful session layer: owns the mutable
        \code{current} bicommand, undo/redo stacks, relational-spec checking,
        rewrite dispatch, invariant suggestion, and export.  Calls into
        \code{http_server.ml} for transport and \code{ui.ml} for HTML; contains
        no socket code itself.
  \item[\code{http_server.ml}] the HTTP transport layer: socket accept loop,
        request-line parsing, response formatting, and graceful SIGINT/SIGTERM
        shutdown.  Has no knowledge of alignments or rewrites.
  \item[\code{ui.ml}] pure HTML/JS rendering: the single-page interactive
        application (\code{build_interactive_html}), served at \code{/ui}.  No
        socket or session logic.
\end{description}

Dependencies flow one way: \code{interactive.ml} uses \code{auto.ml},
\code{align.ml}, \code{rewrites.ml}, \code{typing.ml}, \code{http_server.ml},
and \code{ui.ml}; \code{http_server.ml} and \code{ui.ml} are independent of
each other and of \code{interactive.ml}; \code{align.ml} never refers to
\code{interactive.ml}.




% ============================================================
\section{References}
% ============================================================

\begin{thebibliography}{9}

\bibitem{banerjee-toplas-2022}
A.~Banerjee, R.~Nagasamudram, D.~A. Naumann, and M.~Nikouei.
\newblock A relational program logic with data abstraction and dynamic framing.
\newblock \emph{ACM Transactions on Programming Languages and Systems}, 44(4):25:1--25:136, 2023.
\newblock Issue date: December 2022.
\newblock \url{https://doi.org/10.1145/3551497}.

\bibitem{antonopoulose-2023-popl}
T.~Antonopoulos, E.~Koskinen, T.~C. Le, R.~Nagasamudram, D.~A. Naumann, and M.~Ngo.
\newblock An algebra of alignment for relational verification.
\newblock \emph{Proceedings of the ACM on Programming Languages}, 7(POPL), Article 20, January 2023, 31 pages.
\newblock \url{https://doi.org/10.1145/3571213}.

\end{thebibliography}

\end{document}
